File-based localization of numerical perturbations in data analysis pipelines

### 2020-02-28 paper.tex

The steps implemented in \toolname to compare pipeline executions in
different conditions are outlined in Figure~\ref{fig:overview-tool}. First,
in Condition 1, a directed, rooted, bipartite provenance graph is built for
each subject, where nodes represent application processes or files, and
edges represent read or write file accesses (Step 1 in Figure
Figure~\ref{fig:overview-tool}). \emph{Transient} files are then identified as
 files that are either deleted during the execution, or modified by
multiple processes (Step 2). \toolname re-runs the pipeline to save these files
right after they are produced, resulting in a provenance \emph{DAG}
(Directed Acyclic Graph) in which file nodes have a single parent (in-degree
of 1) (Step 3). This DAG is then compared to the one produced from
Condition 2, in a step-by-step process that prevents the propagation of differences.
\toolname finally produces a labeled graph that identifies the
non-reproducible processes of the pipeline.

@@

  \todo{review and summarize assumptions:
  * No concurrent writes from multiple processes
  * A given process cannot modify the bytes it wrote to a file
  But we support non-concurrent writes by different processes. This is common in neuroimaging.
  }

\subsection{Recording of provenance graphs}

We use the \reprozip tool~\cite{rampin2016reprozip}
to capture: (1) the set of processes created by the
pipeline, using the \texttt{clone()} or \texttt{fork()} system call, and
(2) the set of files read, written and executed by each process, including
temporary files. \reprozip collects this information through the
\texttt{ptrace()} system call, with no required instrumentation of the pipeline.
\toolname then reconstructs a provenance graph by creating process and file
nodes from the \reprozip trace, and by adding directed edges corresponding
to file reads and file writes (Figure~\ref{fig:simple_script}).

@@

Provenance graphs may be subject-dependent, which justifies iterating
\toolname on multiple subjects. For instance, data acquisitions may be
repeated a different number of times in each subject, for noise-reduction
purposes. Some of these differences can be neglected, for instance when a
data decompression step is present at the beginning of the execution for
some subjects only, but other differences cannot be ignored, for instance
when complete different processing paths are used in different subjects.

@@

\subsection{Capture of transient files}

\toolname captures temporary files in Condition 1 by replacing every
process P by a wrapper that first calls P and then saves the produced
temporary files to a read-only directory. This process replacement is done
without pipeline instrumentation, by pre-pending to the PATH environment
variable a directory containing a wrapper script named as the executable
called by the process (see Listing~\ref{listing:modify-script}).

% The temporary files that are removed by P 
% itself won't be saved, but this is not a problem since these files do not play any role in 
% the subsequent steps of the pipeline.

\begin{listing}
  \begin{minted}[frame=single,
    framesep=3mm,
    linenos=false,
    xleftmargin=0pt,
    tabsize=4]{python}
  #!/usr/bin/env python
  import os.path as op
  import shutil
  import pipes

  # Run the original executable using the same arguments
  executed_file = sys.argv[0]
  exec_proc = op.join('original_scripts', executed_file)
  exec_argv = " ".join(map(pipes.quote, sys.argv[1:]))
  exec_cmd = exec_proc + exec_argv
  subprocess.Popen(exec_cmd, shell=True).communicate()

  # Load list of captured process to be checked
  with open('path/to/p_list/', 'r') as p_list:
      process_list = json.load(p_list)

  if exec_cmd in process_list.keys():
      # Save output files to read-only directory
      p_files = process_list[exec_cmd]
      shutil.copy(p_files, '/read-only/dir')

  \end{minted}
    \caption{Wrapper script used to save temporary and modified files.}
    \label{listing:modify-script}
  \end{listing}

Files written by multiple processes are disambiguated as follows. For a
 file F written by the processes in \textbf{P} = \{$P_{1}$, \ldots,
 $P_{n}$\}, \toolname first checks that processes in \textbf{P} do not
 write concurrently to F, which would violate our assumptions. Then, it
 replaces every process $P_{i}$ by a \texttt{PATH}-based wrapper that first
 calls $P_{i}$ and then saves F to a read-only directory. In this way, the
 multiple versions of F are preserved for comparison. \toolname finally
 updates the provenance graph accordingly, so that all files in the graph
 have an in-degree of 1. This operation also makes the provenance graph
 acyclic, since we assumed that a process cannot modify a file once it was
 written \todo{check/prove that}.

@@

\subsection{Process labelling}

Once all the files have been captured in Condition 1, including transient
files, \toolname labels pipeline processes through a step-by-step execution
in Condition 2. Each process in Condition 2 is run and its output files are
compared to the ones produced in Condition 1. If no differences are found,
the process is marked as reproducible. Otherwise, the process is marked as
irrreproducible and the output files produced in Condition 1 are copied to
Condition 2, to ensure that differences do not propagate further in the
pipeline. Processes are instrumented transparently, through a modification
of the \texttt{PATH} variable similar to the one described previously.

By default, differences in output files are identified by comparing file
checksums. Other comparison functions can also be defined for specific file
types, for instance to ignore sections of the file containing timestamps or
other execution-specific information.

Finally, \toolname creates a labeled provenance graph, 
highlighting non-reproducible processes.

>>>>

The steps implemented in \toolname to compare pipeline executions in
2 conditions are outlined in Figure~\ref{fig:overview-tool}. First,
in Condition 1, a directed, rooted, bipartite provenance graph is built for
each subject, where nodes represent application processes or files, and
edges represent read or write file accesses (Step 1 in Figure~\ref{fig:overview-tool}). \emph{Transient} files are then identified as
 files that are either deleted during the execution, or modified by
multiple processes (Step 2). \toolname re-runs the pipeline to save these files
right after they are produced, resulting in a provenance \emph{DAG}
(Directed Acyclic Graph) in which file nodes have a single parent (in-degree
of 1) (Step 3). This DAG is then compared to the one produced from
Condition 2, in a step-by-step process that prevents the propagation of differences (step 4).
\toolname finally produces a labeled graph that identifies the
non-reproducible processes of the pipeline.

@@

  \todo{review and summarize assumptions:
  * No concurrent writes from multiple processes
  * A given process cannot modify the bytes it wrote to a file
  * Process are command lines
  But we support non-concurrent writes by different processes. This is common in neuroimaging.
  }

\subsection{Recording provenance graphs}

We use the \reprozip tool~\cite{rampin2016reprozip}
to capture: (1) the set of processes created by the
pipeline, and
(2) the set of files read and written by each process, including
temporary files. \reprozip collects this information through the
\texttt{ptrace()} system call, with no required instrumentation of the pipeline.
\toolname then reconstructs a provenance graph by creating process and file
nodes from the \reprozip trace, and by adding directed edges corresponding
to file reads and file writes (Figure~\ref{fig:simple_script}).

@@

Provenance graphs may be subject-dependent, which justifies the processing
of multiple subjects. For instance, data acquisitions may be
repeated a different number of times in each subject, for noise-reduction
purposes. Some of these differences can be neglected, for instance when a
data decompression step is present at the beginning of the execution for
some subjects only, but other differences cannot, for instance
when complete different processing paths are used in different subjects.

@@

\subsection{Capturing transient files}

\toolname captures temporary files in Condition 1 by replacing every
process P by a wrapper that first calls P and then saves the produced
temporary files to a read-only directory. This process replacement is done
without pipeline instrumentation, by pre-pending to the \texttt{PATH} environment
variable a directory containing a wrapper script named as the executable
called by P.

Files written by multiple processes are disambiguated as follows. For a
 file F written by the processes in \textbf{P} = \{$P_{1}$, \ldots,
 $P_{n}$\}, \toolname first checks that processes in \textbf{P} do not
 write concurrently to F, which would violate our assumptions. Then, it
 replaces every process $P_{i}$ by a \texttt{PATH}-based wrapper that first
 calls $P_{i}$ and then saves F to a read-only directory. In this way,
 successive versions of F are preserved for comparison. \toolname finally
 updates the provenance graph accordingly, so that all files in the graph
 have an in-degree of 1. This operation also makes the provenance graph
 acyclic, since we assumed that a process can only write a single time to a file.

@@

\subsection{Labelling processes}

Once all the files have been captured in Condition 1, including transient
ones, \toolname labels pipeline processes through a step-by-step execution
in Condition 2. Each process in Condition 2 is run and its output files are
compared to the ones produced in Condition 1. If no differences are found,
the process is marked as reproducible. Otherwise, the process is marked as
non-reproducible and the output files produced in Condition 1 are copied to
Condition 2, to ensure that differences do not propagate further in the
pipeline. Processes are instrumented transparently, through a modification
of the \texttt{PATH} variable similar to the one described previously. By
default, differences in output files are identified by comparing file
checksums. Other comparison functions can also be defined for specific file
types, for instance to ignore sections of the file containing timestamps or
other execution-specific information. Finally, \toolname creates a labeled
provenance graph, highlighting non-reproducible processes.
